---
title: "SI 618 Winter 2017 Homework 8"
author: "Haoyou Liu"
date: "March 22, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(DBI)
library(RSQLite)
library(data.table)
library(ggplot2)
setwd("C:/Users/TF/Desktop/course/SI618/hw8")
```

## Getting Data from SQLite Database

In the data preparation step, a SQLite database has been created and populated with vehicle data. Now the data in the database is retrieved using R package DBI and RSQLite and stored in a data frame named vehicles. Here are the first 10 rows of the data frame, and the summary.

```{r echo=FALSE}
sqlite <- dbDriver("SQLite")
vehiclesdb <- dbConnect(sqlite, "vehicles.db")
vehicles <- dbGetQuery(vehiclesdb, "select * from vehicles")
head(vehicles, n=10)
summary(vehicles)
```

## Converting to Factors

To make downstream analysis easier, we convert the data in columns vehicles$make, vehicles$VClass, vehicles$cylinders, and vehicles$trany into factors. Here is the summary of the data frame after the conversion.

```{r echo=FALSE}
vehicles$make <- as.factor(vehicles$make)
vehicles$vclass <- as.factor(vehicles$vclass)
vehicles$cylinders <- as.factor(vehicles$cylinders)
vehicles$trany <- as.factor(vehicles$trany)
summary(vehicles)
```

## Filter Down Data

We will filter down the data such that only ‘VClass’ with more than 40 vehicles are kept. Here is the summary of the data frame after this subsetting step.

```{r echo=FALSE}
vehicles <- data.table(vehicles)
vehicles <- vehicles[, vclass.ct := .N, by = vclass]
vehicles <- subset(vehicles, vclass.ct > 40)
print(vehicles)
summary(vehicles)
```

## Fuel Economy of Vehicles of Different Makes

For each vehicle class in filtered down data, we plot the mean combined MPG (average of data in vehicles$comb08) for each vehicle maker every year. And then, we compute the mean combined MPG in all years for each vehicle maker, and plot it. Both charts are created with ggplot(). Note how the vehicle makers are ranked in the second plot. Use fig.width=16. To suppress messages from ggplot regarding groups with only one observation, set warning=FALSE, message=FALSE.

```{r echo=FALSE, fig.width = 16}
class = unique(vehicles$vclass)
for (vc in class) {
  sub <- subset(vehicles, vclass==vc)
  data1 <- sub[, mpg := mean(combo08), by = c("year", "make")]
  lineplot <- ggplot(data1, aes(year, mpg)) + geom_line(aes(colour = make)) + guides(col = guide_legend(ncol=2)) + ggtitle(vc) + labs(x = "Year", y = "Mean combined MPG")
  print(lineplot)

  data2 <- sub[, c("make", "combo08")][, mpg := mean(combo08), by = make][, combo08 := NULL]
  data2 <- unique(data2)
  barplot = ggplot(data2, aes(x = reorder(make, mpg, function(x) -x))) + aes(y = mpg) + geom_bar(stat = "identity", alpha = 0.8) + ggtitle(vc) + labs(x = "Make", y = "Mean combined MPG in All Years") + theme(axis.text.x = element_text(angle = 90,hjust = 1))
  print(barplot)
}
```